RegressionPvNP

### 2017-07-19 RegressionPvNP.Rmd
...

```{r setup, include=FALSE}
library(biomaRt)
library(ggplot2)
library(ggfortify)
library(GGally)
library(plyr)
library(pander)
library(reshape2)
library(gridExtra)
library(GenomicRanges)
library(glmnet)
library(CoxHD)
library(mice)

@@

includeDemo = FALSE

@@

## Patient info file
patient.file = grep('All_patient_info.txt', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.txt', data.files, value=T)

patient.info = read.patient.info(patient.file, demo.file, set=dataset)
patient.info$Patient = gsub("/", "_", patient.info$Patient)
head(patient.info)

validation.patient.info = read.patient.info(patient.file, demo.file, set='Test')
validation.patient.info = arrange(validation.patient.info, Status, Patient, Endoscopy.Year, Pathology)
sum.validation.info = summarise.patient.info(validation.patient.info)

patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

@@

sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now

@@

pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

pander(sum.validation.info[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
```

@@

```{r labelsPNP, echo=F, message=F, include=F}
## binomial: dysplasia 1, BE 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

pts = do.call(rbind, lapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  cbind(df$info[,c('Patient','Samplename')])
}))
rownames(pts) = 1:nrow(pts)

@@

validation.df = t(mergedDfValidation[, grep('^D', colnames(mergedDfValidation))])
dim(validation.df)

@@

  ## Some sort of complexity score
  #plot(unlist(complexity.measures), col=sum.patient.data$Status)
  # t.test(
  # complexity.measures[subset(sum.patient.data, Status == 'P')$Patient],
  # complexity.measures[subset(sum.patient.data, Status == 'NP')$Patient]
  # )
  score.cx <-function(pd) {
  #  print(pd$info)
    len.var = (var(pd$seg.vals$end.pos - pd$seg.vals$start.pos))/1e12
    extreme.values = max( apply(as.matrix(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))]), 2, function(col) {
      length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
    }) )

  complexity.measures = sapply(patient.data, score.cx)

  dysplasia.df = cbind(dysplasia.df, 'cx.measure'=NA)
  for (sample in rownames(dysplasia.df)) {
    dysplasia.df[sample, 'cx.measure'] = complexity.measures[subset(patient.info, Samplename == sample)$Patient]
  }
  # norm to 1 to keep within the magnitude of the genomic data
  cx = dysplasia.df[,'cx.measure'] 
  dysplasia.df[,'cx.measure'] = (log(cx)/log(max(cx)))
  #dysplasia.df[,'cx.measure'] = ((cx - min(cx))/(max(cx)-min(cx)))

  complexity.measures = sapply(validation.patient.data, score.cx)

  validation.df = cbind(validation.df, 'cx.measure'=NA)
  for (sample in rownames(validation.df)) {
    validation.df[sample, 'cx.measure'] = complexity.measures[subset(validation.patient.info, Samplename == sample)$Patient]
  }
  # norm to 1 to keep within the magnitude of the genomic data
  cx = validation.df[,'cx.measure'] 
  validation.df[,'cx.measure'] = (log(cx)/log(max(cx)))

@@

create.patient.sets<-function(pts, n, splits, minR=0.2) {
  # This function just makes sure the sets don't become too unbalanced with regards to the labels.
  check.sets<-function(df, grpCol, min) {
    sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
    while ( (length(which(sets/rowSums(sets) < minR) ) >= 2 | length(which(sets/rowSums(sets) == 0)) > 0) ) {
      #print(sets/rowSums(sets))
      s = sample(rep(seq(5), length = length(unique(df$Patient))))
      df2 = merge(df, cbind('Patient'=unique(df$Patient), 'tmpgrp'=s), by="Patient")
      df[[grpCol]] = df2$tmpgrp
      sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
    }
    return(df[[grpCol]])
  }

  s = sample(rep(seq(splits), length = length(unique(pts$Patient))))
  patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")  
  colnames(patients)[3] = c('fold.1')
  patients$fold.1 = check.sets(patients, 'fold.1', minR)

  for (i in 2:n) {
    s = sample(rep(seq(5), length = length(unique(pts$Patient))))
    patients = merge(patients, cbind('Patient'=unique(patients$Patient), 'group'=s), by="Patient")  
    foldcol = grep('group',colnames(patients))
    colnames(patients)[foldcol] = paste('fold',i,sep='.')
    patients[[paste('fold',i,sep='.')]] = check.sets(patients, paste('fold',i,sep='.'))
  }

@@

    cv.class = matrix(nrow=splits, ncol=length(lambda))
    deviance = matrix(nrow=splits, ncol=length(lambda))
    for (i in 1:splits) { 
      message(paste(i, "split"))
      test.rows = which(rownames(x) %in% tpts[which(tpts[,setCol] == i), 'Samplename'])
      test = x[test.rows,]
      training = x[-test.rows,]
      # pre-spec lambda seq
      fitCV <- glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', ...) 
      # autoplot(fit) + theme(legend.position="none")

@@

```{r xvalpt, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
nl = 1000

coefs = list(); plots = list(); performance.at.1se = list()
folds = 10; splits = 5 
alpha.values = c(0,0.5,0.7,0.8,0.9,1)

@@

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=T) # all patients
    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

    l = fit0$lambda
    if (a == 0) {
      l = sort(c(fit0$lambda, seq(exp(-5), exp(-10), -1e-6),
            seq(exp(-10), exp(-15), -1e-8),
            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
    }
    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=T)

@@

    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, lambda.1se == T)
  }
  save(plots, coefs, performance.at.1se, dysplasia.df, file=file)
}

@@

@@

```{r, warning=F, message=F}
#topF = names(which((coef.stable[['1']]) >= 0.75))
topF = names(coef.stable[[select.alpha]])

@@

# Gain or loss not correlated in prog or non
genomeCols = grep('^\\d+:', colnames(dysplasia.df), value=T)
gl = cbind.data.frame(
  'P.gain' = apply(features.in.prog, 2, function(x) sum( as.integer(x >= 1.1) ) ), 
  'P.loss' = apply(features.in.prog, 2, function(x) sum( as.integer(x <= 0.9) ) ),
  'NP.gain' = apply(features.nonprog, 2, function(x) sum( as.integer(x >= 1.1) ) ),
  'NP.loss' = apply(features.nonprog, 2, function(x) sum( as.integer(x <= 0.9) ) ) )

allgl = cbind.data.frame(
  'ALL.P.gain' = apply(dysplasia.df[names(which(labels == 1)), genomeCols], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
  'ALL.P.loss' = apply(dysplasia.df[names(which(labels == 1)),genomeCols], 2, function(x) sum( as.integer(x <= 0.9) ) ),
  'ALL.NP.gain' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
  'ALL.NP.loss' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x <= 0.9) ) ) )

@@

# Gains/losses enriched in the top features?  Appears so
(fisher.test(rbind(colSums(gl[,c('P.gain','P.loss')]),colSums(allgl[-feat,c('ALL.P.gain','ALL.P.loss')])  )) )

@@

topFSegments = do.call(rbind,  strsplit( topF[grep('^\\d+:', topF)], ':|-' ))
colnames(topFSegments) = c('chr','start','end')
topFSegments = as.data.frame(apply(topFSegments, 2, as.numeric))
topFSegments$chr = factor(topFSegments$chr, levels=chromInfo$chr)
topFSegments = arrange(topFSegments, chr, start)

@@

grid.arrange(ggplot(topFRegions, aes(reorder(chr, -adj.info.content), totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
  labs(y='Feature bp/total chr length', title='Ratio of covered base pairs', x='chr') +
  geom_bar(stat='identity') +  geom_text(aes(label=round(totalbp/length, 2)), vjust=-0.5) +
  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom'), 
  bottom='Ratio of base pair coverage per chromosome in the top non-zero features\n selected from the model. Chromosomes are sorted\nand binned by their gene (e.g. information) content rather than by overall length.')

@@

  # Not related to the length of the segment
  pander(with(topFSegments, cor.test(genes, (end-start))), caption="Number of genes per segment is not related to the segment length.")
}, error = function(e) {
  warning(paste("Failed to connect to Biomart service", e))
})
```

#### Correlations

```{r echo=F, message=F, warning=F}
library(Hmisc)

@@

nrow(rc)
head(rc)

@@

head(rc)

@@

rc = arrange(rc, row.chr, col.chr, row.start, col.start)
head(rc)

@@

head(rc1)

@@

```{r echo=F, warning=F, message=F,eval=T, fig.height=8, fig.width=8}
hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})
ch = as.data.frame(hazards)
ch$coef = coefs[[select.alpha]][,1]
ch$label = rownames(ch)
library(ggrepel)
ggplot(ch, aes(coef, hazards, col=hazards > quantile(hazards)[['75%']])) + 
  #xlim(min(coefGR$coef)-5, max(coefGR$coef)+10) +
  geom_point(alpha=0.5) + 
  geom_text_repel(aes(x=coef+10, label=ifelse(hazards > quantile(hazards)[['75%']], label, '')),show.legend=F) +
  labs(title='Top features coef vs hazard') + theme(legend.position = 'bottom')
```

```{r echo=F, warning=F, message=F,eval=T, fig.height=12, fig.width=8}
ch = arrange(ch, -hazards)

ch$hazards = round(ch$hazards, 2)
ch$coef = round(ch$coef, 2)

@@

  reg = cbind.data.frame( 'value'=dysplasia.df[ , r ], labels)
  reg$labels = factor(reg$labels)
  reg$labels = revalue(reg$labels, c("1"="P", "0"="NP"))

@@

  reg$pval = tt$p.value
  reg$region = r
  reg$hazard = x[['hazards']]
  reg$coef = x[['coef']]
  return(reg)
}

plot.reg<-function(x) {
  reg = melt.by.reg(x)
  pv = ifelse(tt$p.value < .001, '< .001', format.pval(unique(reg$p.value)))

#  ggplot(reg, aes(value, fill=labels) ) + geom_histogram(bins=50, color='grey') + 
#    labs(x='seg. value', title=r, subtitle=paste('haz=',x[['hazards']], ' coef=', x[['coef']], ' MW pval: ',pv,sep=''))
  ggplot(reg, aes(x=labels, y=value, group=labels, fill=labels) ) + geom_jitter(color='grey', show.legend = F) + geom_violin(show.legend = F, alpha=0.5) +  ylim(0.75, 1.2) +
    geom_hline(yintercept = median(reg$value) + sd(reg$value)) + geom_hline(yintercept = median(reg$value) - sd(reg$value)) +
    labs(x='', y='seg. value', title=r, subtitle=paste('haz=',x[['hazards']], ' coef=',x[['coef']], '\nMW pval ',pv,sep=''))
}

regions = apply(ch, 1, melt.by.reg)

do.call(grid.arrange, c(lapply(regions, function(reg) {
  se = 2
  x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value >= median(value) + sd(value)*se))),
                               'loss' = length(which(with(subset(reg, labels == 'P'), value <= median(value) - sd(value)*se))) ),

                  'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value >= median(value) + sd(value)*se))),
                                'loss' = length(which(with(subset(reg, labels == 'NP'), value <= median(value) - sd(value)*se))) ) )
  # x = rbind.data.frame('P'= cbind( 'high' = length(which(with(subset(reg, labels == 'P'), value >= 1.1))),
  #                                  'low' = length(which(with(subset(reg, labels == 'P'), value <= 0.9))) ),
  # 
  #                 'NP'= cbind( 'high' = length(which(with(subset(reg, labels == 'NP'), value >= 1.1))),
  #                                  'low' = length(which(with(subset(reg, labels == 'NP'), value <= 0.9))) ))

  cols = c('indianred3','slateblue2')
  if ( as.numeric(reg$coef[1]) < 0) cols = c('plum', 'skyblue3')
  #ggplot(melt(as.matrix(x)), aes(Var1, value, fill=Var2) ) + geom_bar(stat='identity', position='dodge') + labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)

  ggplot(melt(as.matrix(rowSums(x))), aes(Var1, value, fill=Var1)) + geom_bar(stat='identity', show.legend = F)+ labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)

})))

@@

```{r warning=F, message=F, fig.height=8, fig.width=8}
run.cv<-function(x, y, alphas=c(0,0.2,0.5,1), ...) {
  plots = list(); coefs = list(); performance.at.1se = list()
  for (a in alphas) {
    fitg <- glmnet(x, y, alpha=a, nlambda=nl, family='binomial') # all patients

@@

    cvp = crossvalidate.by.patient(x, y, fit=fitg, lambda=l, a=a, ...)
    plots[[as.character(a)]] = arrangeGrob(cvp$plot+ggtitle('Classification'), cvp$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

@@

  for (a in alpha.values) {
    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl) # all patients

>>>>

```{r setup, include=FALSE}
library(biomaRt)
library(ggplot2)
library(ggrepel)
library(ggfortify)
library(GGally)
library(plyr)
library(dplyr)
library(pander)
library(reshape2)
library(gridExtra)
library(GenomicRanges)
library(glmnet)
library(CoxHD)
library(mice)

@@

includeDemo = FALSE # demographics

@@

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

patient.info = read.patient.info(patient.file, demo.file, set=dataset)$info
head(patient.info)

validation.patient.info = read.patient.info(patient.file, demo.file, set='Test')$info
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.validation.info = summarise.patient.info(validation.patient.info)

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

@@

sum.patient.data = as.data.frame(subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))) ## For now

@@

pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
```

@@

```{r labelsPNP, echo=F, message=F, include=T}
## binomial: dysplasia 1, BE 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

pts = do.call(rbind, lapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  cbind(df$info[,c('Hospital.Research.ID','Samplename')])
}))
rownames(pts) = 1:nrow(pts)

@@

unit.var <- function(x) {
  if (sd(x) == 0) return(x)
  return((x-mean(x))/sd(x) )
}

#dysplasia.df = apply(dysplasia.df, 2, function(x) x-median(x))
dysplasia.df = apply(dysplasia.df, 2, unit.var)

@@

  ## Some sort of complexity score
  #plot(unlist(complexity.measures), col=sum.patient.data$Status)
  # t.test(
  # complexity.measures[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID],
  # complexity.measures[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID]
  # )

  score.cx <-function(pd) {
    samples = grep('^D', colnames(pd$seg.vals), value=T)
    len.var = var(pd$seg.vals$end.pos - pd$seg.vals$start.pos) # length won't vary between samples

    #extremes = apply(as.matrix(pd$seg.vals[,samples]), 2, function(col) {
    #  length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
    #})

    num.var.reg = vector(mode='integer', length=length(samples))
    if (!is.null(pd$variable.regions) && nrow(pd$variable.regions) > 0) {
      num.var.reg = apply(as.matrix(pd$variable.regions[,samples]), 2, function(col) {
        length(which(col >= 1.1 | col <= 0.9))
      })
    }

    sample.var = apply(as.matrix(pd$seg.vals[, samples]), 2, var)
    t(cbind(len.var, num.var.reg, sample.var))
  }

  complexity.measures = lapply(patient.data, score.cx)

  dysplasia.df = cbind(dysplasia.df, 'segment.length'=0, 'num.var.regions'=0, 'sample.variance'=0)
  for (pt in names(complexity.measures)) {
    for (sample in colnames(complexity.measures[[pt]])) {
      dysplasia.df[sample,c('segment.length', 'num.var.regions', 'sample.variance')] = as.numeric(complexity.measures[[pt]][,sample])
    }
    #for (sample in rownames(dysplasia.df)) {
      #dysplasia.df[sample, ] = 
        #complexity.measures[subset(patient.info, Samplename == sample)$Hospital.Research.ID]
    #}
  }
  # norm to 1 to keep within the magnitude of the genomic data
  scale.to.1<-function(x) {
    (x-min(x))/(max(x)-min(x))
  }

  dysplasia.df[,c('segment.length', 'num.var.regions', 'sample.variance')] = apply(dysplasia.df[,c('segment.length', 'num.var.regions', 'sample.variance')], 2, scale)

  #cx = dysplasia.df[,'cx.measure'] 
  #dysplasia.df[,'cx.measure'] = (log(cx)/log(max(cx)))
  #dysplasia.df[,'cx.measure'] = dysplasia.df[,'cx.measure'] - median(dysplasia.df[,'cx.measure'])
  #dysplasia.df[,'cx.measure'] = ((cx - min(cx))/(max(cx)-min(cx)))

@@

create.patient.sets<-function(pts, n, splits, minR=0.2) {
  # This function just makes sure the sets don't become too unbalanced with regards to the labels.
  check.sets<-function(df, grpCol, min) {
    sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
    while ( (length(which(sets/rowSums(sets) < minR) ) >= 2 | length(which(sets/rowSums(sets) == 0)) > 0) ) {
      #print(sets/rowSums(sets))
      s = sample(rep(seq(5), length = length(unique(df$Hospital.Research.ID))))
      df2 = merge(df, cbind('Hospital.Research.ID'=unique(df$Hospital.Research.ID), 'tmpgrp'=s), by="Hospital.Research.ID")
      df[[grpCol]] = df2$tmpgrp
      sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
    }
    return(df[[grpCol]])
  }

  s = sample(rep(seq(splits), length = length(unique(pts$Hospital.Research.ID))))
  patients = merge(pts, cbind('Hospital.Research.ID'=unique(pts$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
  colnames(patients)[3] = c('fold.1')
  patients$fold.1 = check.sets(patients, 'fold.1', minR)

  for (i in 2:n) {
    s = sample(rep(seq(5), length = length(unique(pts$Hospital.Research.ID))))
    patients = merge(patients, cbind('Hospital.Research.ID'=unique(patients$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
    foldcol = grep('group',colnames(patients))
    colnames(patients)[foldcol] = paste('fold',i,sep='.')
    patients[[paste('fold',i,sep='.')]] = check.sets(patients, paste('fold',i,sep='.'))
  }

@@

    cv.class = matrix(nrow=splits, ncol=length(lambda))
    deviance = matrix(nrow=splits, ncol=length(lambda))
    for (i in 1:splits) { 
      message(paste(i, "split"))
      test.rows = which(rownames(x) %in% tpts[which(tpts[,setCol] == i), 'Samplename'])
      test = x[test.rows,]
      training = x[-test.rows,]
      # pre-spec lambda seq
      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', ...) 
      # autoplot(fit) + theme(legend.position="none")

@@

```{r xvalpt, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
# dysplasia.df = cbind(dysplasia.df, 'sl.nvr'=0)
# dysplasia.df[, 'sl.nvr'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'num.var.regions'])
# 
# dysplasia.df = cbind(dysplasia.df, 'sl.nvr.sv'=0)
# dysplasia.df[, 'sl.nvr.sv'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'num.var.regions'] + dysplasia.df[, 'sample.variance'])
# 
# dysplasia.df = cbind(dysplasia.df, 'sl.sv'=0)
# dysplasia.df[, 'sl.sv'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'sample.variance'])
# 
# dysplasia.df = cbind(dysplasia.df, 'nvr.sv'=0)
# dysplasia.df[, 'sl.sv'] = scale(dysplasia.df[, "num.var.regions"] + dysplasia.df[, 'sample.variance'])

nl = 1000

coefs = list(); plots = list(); performance.at.1se = list(); models = list();
folds = 10; splits = 5 
alpha.values = c(0, 0.5,0.7,0.8,0.9,1)

@@

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

    l = fit0$lambda
    if (a == 0) {
      l = sort(c(fit0$lambda, seq(exp(-5), exp(-10), -1e-6),
            seq(exp(-10), exp(-15), -1e-8),
            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
    }
    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)

@@

    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, lambda.1se == T)
    models[[as.character(a)]] = fit0
  }
  save(plots, coefs, performance.at.1se, dysplasia.df, models, file=file)
}

@@

```{r echo=F, message=F, warning=F}
featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
```

@@

```{r, warning=F, message=F, eval=T}
#topF = names(which((coef.stable[['1']]) >= 0.75))
topF = names(coef.stable[[select.alpha]])

@@

# Gain or loss not correlated in prog or non
genomeCols = grep('^\\d+:', colnames(dysplasia.df), value=T)
gl = cbind.data.frame(
  'P.gain' = apply(features.in.prog, 2, function(x) sum( as.integer(x > 0) ) ), 
  'P.loss' = apply(features.in.prog, 2, function(x) sum( as.integer(x < 0) ) ),
  'NP.gain' = apply(features.nonprog, 2, function(x) sum( as.integer(x > 0) ) ),
  'NP.loss' = apply(features.nonprog, 2, function(x) sum( as.integer(x < 0) ) ) )

allgl = cbind.data.frame(
  'ALL.P.gain' = apply(dysplasia.df[names(which(labels == 1)), genomeCols], 2, function(x) sum( as.integer(x > 0) ) ), 
  'ALL.P.loss' = apply(dysplasia.df[names(which(labels == 1)),genomeCols], 2, function(x) sum( as.integer(x < 0) ) ),
  'ALL.NP.gain' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x > 0) ) ), 
  'ALL.NP.loss' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x < 0) ) ) )

@@

# Gains/losses enriched in the top features?  Appears not
(fisher.test(rbind(colSums(gl[,c('P.gain','P.loss')]),colSums(allgl[-feat,c('ALL.P.gain','ALL.P.loss')])  )) )

@@

topFSegments = do.call(rbind,  strsplit( topF[grep('^\\d+:', topF)], ':|-' ))
colnames(topFSegments) = c('chr','start','end')
topFSegments = as.data.frame(apply(topFSegments, 2, as.numeric))
topFSegments$chr = factor(topFSegments$chr, levels=droplevels(chromInfo$chr[1:22]), ordered=T)
topFSegments = arrange(topFSegments, chr, start)

@@

grid.arrange(
  #ggplot(topFRegions, aes(reorder(chr, -adj.info.content), totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
  ggplot(topFRegions, aes(chr, totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
    labs(y='Feature bp/total chr length', title='Ratio of covered base pairs', x='chr') +
    geom_bar(stat='identity') +  geom_text(aes(label=round(totalbp/length, 2)), vjust=-0.5) +
    scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom'), 
      bottom='Ratio of base pair coverage per chromosome in the top non-zero features\n selected from the model. Chromosomes are sorted\nand binned by their gene (e.g. information) content rather than by overall length.')

@@

  # Not related to the length of the segment
  pander(with(topFSegments, cor.test(genes, (end-start))), caption="Number of genes per segment may be related to the segment length.")
}, error = function(e) {
  warning(paste("Failed to connect to Biomart service", e))
})
```

```{r echo=F, message=F, warning=F, eval=F}
#### Correlations
library(Hmisc)

@@

#nrow(rc)
#head(rc)

@@

#head(rc)

@@

rc = arrange(rc, row.chr, col.chr, row.start, col.start)
#head(rc)

@@

#head(rc1)

@@

```{r echo=F, warning=F, message=F,eval=T, fig.height=8, fig.width=8}

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})
ch = as.data.frame(hazards)
ch$coef = coefs[[select.alpha]][,1]
ch$label = rownames(ch)
ch$stability = featureStability

ggplot(ch, aes(coef, hazards, col=hazards > quantile(hazards)[['75%']])) + 
  #xlim(min(coefGR$coef)-5, max(coefGR$coef)+10) +
  geom_point(alpha=0.5) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards > quantile(hazards)[['75%']], label, '')),show.legend=F) +
  labs(title='Top features coef vs hazard') + theme(legend.position = 'bottom')
```

```{r echo=F, warning=F, message=F,eval=F, fig.height=12, fig.width=8}
ch = arrange(ch, -stability, -hazards, -coef)
ch$hazards = round(ch$hazards, 3)
ch$coef = round(ch$coef, 3)
write.table(ch[c('label', 'stability','hazards','coef')], row.names = F, quote=F, sep="\t", file='features.txt')

ch = arrange(ch, -hazards)

@@

  reg = cbind.data.frame( 'value'=dysplasia.df[ ,r], labels)
  reg$labels = factor(reg$labels)
  reg$labels = revalue(reg$labels, c("1"="P", "0"="NP"))

@@

  reg$pval = tt$p.value
  reg$region = r
  reg$hazard = as.numeric(x[['hazards']])
  reg$coef = as.numeric(x[['coef']])
  return(reg)
}

plot.reg<-function(x) {
  reg = melt.by.reg(x)

  pv = ifelse(reg$pval[1] < .001, '< .001', format.pval(unique(reg$pval[1])))

#  ggplot(reg, aes(value, fill=labels) ) + geom_histogram(bins=50, color='grey') + 
#    labs(x='seg. value', title=r, subtitle=paste('haz=',x[['hazards']], ' coef=', x[['coef']], ' MW pval: ',pv,sep=''))
  ggplot(reg, aes(x=labels, y=value, group=labels, fill=labels) ) + geom_jitter(color='grey', show.legend = F) + geom_violin(show.legend = F, alpha=0.5) +
    geom_hline(yintercept = median(reg$value) + sd(reg$value)) + geom_hline(yintercept = median(reg$value) - sd(reg$value)) +
    labs(x='', y='seg. value', title=x[['label']], subtitle=paste('haz= ',x[['hazards']], ' coef= ',x[['coef']], '\nMW pval ',pv,sep=''))
}

ch = arrange(ch, -hazards)

regions = apply(ch, 1, melt.by.reg)
names(regions) = ch$label

do.call(grid.arrange, c(lapply(regions[1:15], function(reg) {
  print(reg$region[1])
  cols = c('indianred3','slateblue2')
  if ( as.numeric(reg$coef[1]) < 0) cols = c('green3', 'darkviolet')

  ggplot(reg, aes(labels, value*coef, fill=labels, color=labels)) + 
    geom_jitter(aes(labels, value), width=0.05, color='yellow4', alpha=0.3) +
    geom_jitter(aes(labels, value*coef, color=labels), width=0.05, alpha=0.5) +
    geom_text_repel(data=subset(reg, value >= quantile(reg$value)[['100%']]), aes(labels, value, label=round(value, 3)), color='yellow4') +
    geom_text_repel(data=subset(reg, value >= quantile(reg$value)[['100%']]), aes(labels, value*coef, label=round(value, 3),color=labels)) + geom_violin(alpha=0.2) + 
    scale_fill_manual(values=cols) + scale_color_manual(values=cols) +
    theme(legend.position = 'none') + labs(title=unique(reg$region)[1], x='')
})))

    #geom_violin(aes(labels, value, fill=labels), alpha=0.5) + theme(legend.position = 'none')

#  x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value > 0))),
#                               'loss' = length(which(with(subset(reg, labels == 'P'), value < 0))) ),

#                        'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value > 0))),
#                                'loss' = length(which(with(subset(reg, labels == 'NP'), value < 0))) ) )

    # x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value >= median(value) + sd(value)*se))),
  #                              'loss' = length(which(with(subset(reg, labels == 'P'), value <= median(value) - sd(value)*se))) ),
  # 
  #                 'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value >= median(value) + sd(value)*se))),
  #                               'loss' = length(which(with(subset(reg, labels == 'NP'), value <= median(value) - sd(value)*se))) ) )
  # x = rbind.data.frame('P'= cbind( 'high' = length(which(with(subset(reg, labels == 'P'), value >= 1.1))),
  #                                  'low' = length(which(with(subset(reg, labels == 'P'), value <= 0.9))) ),
  # 
  #                 'NP'= cbind( 'high' = length(which(with(subset(reg, labels == 'NP'), value >= 1.1))),
  #                                  'low' = length(which(with(subset(reg, labels == 'NP'), value <= 0.9))) ))

  #cols = c('indianred3','slateblue2')
  #if ( as.numeric(reg$coef[1]) < 0) cols = c('plum', 'skyblue3')

  #ggplot(melt(as.matrix(x)), aes(Var2, value, fill=Var1)) + facet_grid(~Var1) +
  #  geom_col( show.legend = F) + labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)

#})))

@@

```{r warning=F, message=F, fig.height=8, fig.width=8}
run.cv<-function(x, y, alphas=c(0,0.2,0.5,1), ...) {
  plots = list(); coefs = list(); performance.at.1se = list()
  for (a in alphas) {
    fitg <- glmnet(x, y, alpha=a, nlambda=nl, family='binomial', standardize = F) # all patients

@@

    cvp = crossvalidate.by.patient(x, y, fit=fitg, lambda=l, a=a, pts=sets, standardize=F)
    plots[[as.character(a)]] = arrangeGrob(cvp$plot+ggtitle('Classification'), cvp$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

@@

  for (a in alpha.values) {
    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl, standardize = F) # all patients
